Analyzing Performance Bottleneck in Graph Neural Network Training: An Experimental View

### 2020-07-23 Manuscript/ms.tex
更新完成初稿.

	\begin{abstract}
		Graph neural network has become a research hotspot in the field of artificial intelligence due to state-of-the-art of performance
		achieved in many grpah-related tasks. At the same time, various of graph neural network system has emrged one after another.
		These systems have used their own skills in implementation, but there are very few works to study the performance bottleneck
		on graph neural network training. In contrast to prior works that present characterization of GCNs and select GNNs cover GNN workloads,
		we select typical algorithms by complexity of edge/vertex calculation. We analyze the performance bottleneck by breaking down training
		time, study the affects of hyper-parameters and sampling technologies and find factors of memory usage. We finally put forward some key
		findings for GNN training and hope our work can help GNN system researchers.
	\end{abstract}

>>>>

	\begin{abstract}
		Graph neural network has become a research hotspot in the field of artificial intelligence due to state-of-the-art of performance
		achieved in many graph-related tasks. At the same time, various of graph neural network systems have emerged one after another.
		These systems have used their own skills in implementation, but there are very few works to study the performance bottleneck
		on graph neural network training. In contrast to prior works that present characterization of GCNs and select GNNs cover GNN workloads,
		we select typical algorithms by complexities of edge/vertex calculation. We analyze the performance bottleneck by breaking down training
		time, study the affects of hyper-parameters and sampling technologies and find factors of memory usage. We finally put forward some key
		findings for GNN training and hope our work can help GNN system researchers.
	\end{abstract}

### 2020-07-23 Manuscript/sec2-review-of-graph-neural-network.tex
update
Update section 1.
更新完成初稿.

\section{Review of Graph Neural Networks}

@@

\begin{figure}
	\centering
	\includegraphics[width=1\columnwidth]{figs/illustration/GNN_common_architecture.png}
	\caption{Structure of a typical graph neural network. (a) Demo GNN, (b) Demo graph. The target application is the node classification. The demo GNN has two GNN layers.}
	\label{fig:general_structure_of_gnn}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\columnwidth]{figs/illustration/GNN_Unit.png}
	\caption{Graph neuron of $v_3$ at the GNN layer $l$ with the graph $\mathcal{G}$ in \figurename~\ref{fig:general_structure_of_gnn}(b). $\phi$/$\Sigma$/$\gamma$ are the message/aggregation/vertex update functions in the message-passing model, respecitvely.}
	\label{fig:graph_neuron_structure}
\end{figure}

@@

			AGCN                                                                                                                      &

@@

			GraphSAGE                                                                                                                 &

@@

			GraphSAGE-pool                                                                                                            &

@@

			Neural FPs                                                                                                                &

@@

			SSE                                                                                                                       &

@@

			\textbf{GGNN}                                                                                                             &

@@

			\textbf{GAT}                                                                                                              &

@@

                         & \text{Sub-layer 0:}\\
                         & \boldsymbol{m}^{l,0}_{j,i,(k)} = \exp(LeakyReLU(\textcolor{blue}{\boldsymbol{a}^T} [\hat{\boldsymbol{h}}^{l}_{i,(k)} \parallel \hat{\boldsymbol{h}}^{l}_{j, (k)}]))\\
                         & \text{Sub-layer 1:}\\

@@

                \text{concat: } O(d_{out}) &\\
                \text{average: } O(K * d_{out}) &\\

@@

			\textbf{GaAN}                                                                                                             &

@@

						 & \alpha^l_{j, i, (k)} = \frac {\exp((\textcolor{blue}{W^{l, k}_{xa}} \boldsymbol{h}_j^l + \textcolor{blue}{b^{l, k}_{xa}})^T (\textcolor{blue}{\boldsymbol{W}^{l, k}_{ya}} \boldsymbol{h}_i^l + \textcolor{blue}{\boldsymbol{b}^{l,k}_{ya}}))} {\hat{\boldsymbol{a}}^{l, 0}_{i, (k)}}\\

@@

                        Four sub-layers}\\

@@

                & \text{Preprocessing:}\\
                & \hat{\boldsymbol{h}}^{l}_{i,(k)} = \textcolor{blue}{\boldsymbol{W}^{l,k}} \boldsymbol{h}_i^{l}\\
                & \text{Sub-layer 0:}\\
                & \hat{\boldsymbol{a}}^{l,0}_{i,(k)} = \boldsymbol{s}^{l,0}_{i,(k)}\\
                & \text{Sub-layer 1:}\\

@@

             \text{concat:} O(d_{in}*d_{out}) &\\
             \text{average:} O(K*d_{in}*d_{out}) &\\
             \text{Two sub-layers}&

@@

						& \text{Sub-layer 0:}\\
						& \hat{\boldsymbol{a}}^{l,0}_{i,(k)} = \boldsymbol{s}^{l,0}_{i,(k)}\\
						& \text{Sub-layer 1:}\\

@@

			\makecell[r]{$O(max(K * d_v + d_{in}, 2 * d_{in} + d_m) * d_{out})$\\

>>>>

\section{Review of Graph Neural Networks}
\label{sec:review_of_gnns}

@@

\begin{figure}
	\centering
	\includegraphics[width=1\columnwidth]{figs/illustration/GNN_common_architecture.pdf}
	\caption{Structure of a typical graph neural network. (a) Demo GNN, (b) Demo graph. The target application is the node classification. The demo GNN has two GNN layers.}
	\label{fig:general_structure_of_gnn}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\columnwidth]{figs/illustration/GNN_Unit.pdf}
	\caption{Graph neuron of $v_3$ at the GNN layer $l$ with the graph $\mathcal{G}$ in \figurename~\ref{fig:general_structure_of_gnn}(b). $\phi$/$\Sigma$/$\gamma$ are the message/aggregation/vertex update functions in the message-passing model, respecitvely.}
	\label{fig:graph_neuron_structure}
\end{figure}

@@

			AGCN \cite{li2018_agcn}                                                                                                   &

@@

			GraphSAGE \cite{hamilton2017_graphsage}                                                                                   &

@@

			GraphSAGE-pool \cite{hamilton2017_graphsage}                                                                              &

@@

			Neural FPs  \cite{duvenaud2015_neural_fps}                                                                                &

@@

			SSE \cite{han2018_sse}                                                                                                    &

@@

			\textbf{GGNN}  \cite{li2015_ggnn}                                                                                         &

@@

			\textbf{GAT}   \cite{huang2018_gat}                                                                                       &

@@

						 & \boldsymbol{m}^{l,0}_{j,i,(k)} = \exp(LeakyReLU(\textcolor{blue}{\boldsymbol{a}^T} [\hat{\boldsymbol{h}}^{l}_{i,(k)} \parallel \hat{\boldsymbol{h}}^{l}_{j, (k)}]))                                    \\

@@

					\text{concat: } O(d_{out})      & \\
					\text{average: } O(K * d_{out}) & \\

@@

			\textbf{GaAN}     \cite{zhang2018_gaan}                                                                                   &

@@

						 & \alpha^l_{j, i, (k)} = \frac {\exp((\textcolor{blue}{W^{l, k}_{xa}} \boldsymbol{h}_j^l + \textcolor{blue}{b^{l, k}_{xa}})^T (\textcolor{blue}{\boldsymbol{W}^{l, k}_{ya}} \boldsymbol{h}_i^l + \textcolor{blue}{\boldsymbol{b}^{l,k}_{ya}}))} {\hat{\boldsymbol{a}}^{l, 0}_{i, (k)}} \\

@@

			Four sub-layers}                                                                                                            \\

@@

						 & \text{Preprocessing:}                                                                          \\
						 & \hat{\boldsymbol{h}}^{l}_{i,(k)} = \textcolor{blue}{\boldsymbol{W}^{l,k}} \boldsymbol{h}_i^{l} \\
						 & \hat{\boldsymbol{a}}^{l,0}_{i,(k)} = \boldsymbol{s}^{l,0}_{i,(k)}                              \\

@@

					\text{concat:} O(d_{in}*d_{out})    & \\
					\text{average:} O(K*d_{in}*d_{out}) & \\

@@

						 & \hat{\boldsymbol{a}}^{l,0}_{i,(k)} = \boldsymbol{s}^{l,0}_{i,(k)}                                                                                                                              \\
						 & \text{Sub-layer 1, 2, 3:}                                                                                                                                                                      \\

@@

			\makecell[r]{$O(max(K * d_v + d_{in}, 2 * d_{in} + d_m) * d_{out})$                                                                                                                                                \\

### 2020-07-23 Manuscript/sec4-experiments.tex
更新完成初稿.

\begin{figure}
    \centering
    \subfloat[GCN]{\includegraphics[height=3cm]{figs/experiments/exp_hyperparameter_on_memory_usage_gcn.png}}
    \subfloat[GGNN]{\includegraphics[height=3cm]{figs/experiments/exp_hyperparameter_on_memory_usage_ggnn.png}}\\
    \subfloat[GAT]{\includegraphics[height=3cm]{figs/experiments/exp_hyperparameter_on_memory_usage_gat.png}}\\
    \subfloat[GaAN]{\includegraphics[height=3cm]{figs/experiments/exp_hyperparameter_on_memory_usage_gaan.png}}
    \caption{Effects of hyper-parameters on the peak GPU memory usage during the training, excluding the memory used by the dataset and the model parameters.}
    \label{fig:exp_hyperparameter_memory_usage}
\end{figure}

@@

\begin{figure}
    \centering
    \subfloat[GCN]{\includegraphics[height=4cm]{figs/experiments/exp_avg_degree_on_vertex_edge_cal_time_gcn.png}}
    \subfloat[GGNN]{\includegraphics[height=4cm]{figs/experiments/exp_avg_degree_on_vertex_edge_cal_time_ggnn.png}}\\
    \subfloat[GAT]{\includegraphics[height=4cm]{figs/experiments/exp_avg_degree_on_vertex_edge_cal_time_gat.png}}
    \subfloat[GaAN]{\includegraphics[height=4cm]{figs/experiments/exp_avg_degree_on_vertex_edge_cal_time_gaan.png}}
    \caption{Effects of the average degree on the time proportion of the vertex/edge calculation. Graphs were generated with the R-MAT generator by fixing the number of vertices as 50,000. }
    \label{fig:exp_avg_degree_on_vertex_edge_cal_time}
\end{figure}

@@

\begin{figure}
    \centering
    \subfloat[GCN]{\includegraphics[height=4cm]{figs/experiments/exp_edge_calc_decomposition_gcn.png}}
    \subfloat[GGNN]{\includegraphics[height=4cm]{figs/experiments/exp_edge_calc_decomposition_ggnn.png}}\\
    \subfloat[GAT]{\includegraphics[height=4cm]{figs/experiments/exp_edge_calc_decomposition_gat.png}}
    \subfloat[GaAN]{\includegraphics[height=4cm]{figs/experiments/exp_edge_calc_decomposition_gaan.png}}
    \caption{Training time breakdown of the edge calculation stage (including both GNN layers).}
    \label{fig:exp_edge_calc_decomposition}
\end{figure}

>>>>

\begin{figure}
    \centering
    \subfloat[GCN]{\includegraphics[height=3cm]{figs/experiments/exp_hyperparameter_on_memory_usage_gcn.pdf}}
    \subfloat[GGNN]{\includegraphics[height=3cm]{figs/experiments/exp_hyperparameter_on_memory_usage_ggnn.pdf}}\\
    \subfloat[GAT]{\includegraphics[height=3cm]{figs/experiments/exp_hyperparameter_on_memory_usage_gat.pdf}}\\
    \subfloat[GaAN]{\includegraphics[height=3cm]{figs/experiments/exp_hyperparameter_on_memory_usage_gaan.pdf}}
    \caption{Effects of hyper-parameters on the peak GPU memory usage during the training, excluding the memory used by the dataset and the model parameters.}
    \label{fig:exp_hyperparameter_memory_usage}
\end{figure}

@@

\begin{figure}
    \centering
    \subfloat[GCN]{\includegraphics[height=4cm]{figs/experiments/exp_avg_degree_on_vertex_edge_cal_time_gcn.pdf}}
    \subfloat[GGNN]{\includegraphics[height=4cm]{figs/experiments/exp_avg_degree_on_vertex_edge_cal_time_ggnn.pdf}}\\
    \subfloat[GAT]{\includegraphics[height=4cm]{figs/experiments/exp_avg_degree_on_vertex_edge_cal_time_gat.pdf}}
    \subfloat[GaAN]{\includegraphics[height=4cm]{figs/experiments/exp_avg_degree_on_vertex_edge_cal_time_gaan.pdf}}
    \caption{Effects of the average degree on the time proportion of the vertex/edge calculation. Graphs were generated with the R-MAT generator by fixing the number of vertices as 50,000. }
    \label{fig:exp_avg_degree_on_vertex_edge_cal_time}
\end{figure}

@@

\begin{figure}
    \centering
    \subfloat[GCN]{\includegraphics[height=4cm]{figs/experiments/exp_edge_calc_decomposition_gcn.pdf}}
    \subfloat[GGNN]{\includegraphics[height=4cm]{figs/experiments/exp_edge_calc_decomposition_ggnn.pdf}}\\
    \subfloat[GAT]{\includegraphics[height=4cm]{figs/experiments/exp_edge_calc_decomposition_gat.pdf}}
    \subfloat[GaAN]{\includegraphics[height=4cm]{figs/experiments/exp_edge_calc_decomposition_gaan.pdf}}
    \caption{Training time breakdown of the edge calculation stage (including both GNN layers).}
    \label{fig:exp_edge_calc_decomposition}
\end{figure}
